---
output: github_document
always_allow_html: true
editor_options: 
  markdown: 
    wrap: 72
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  message = FALSE,
  warning = FALSE,
  fig.retina = 2,
  fig.align = 'center'
)
```

# ploswater

<!-- badges: start -->

[![License: CC BY
4.0](https://img.shields.io/badge/License-CC_BY_4.0-lightgrey.svg)](https://creativecommons.org/licenses/by/4.0/)
[![R-CMD-check](https://github.com/openwashdata/ploswater/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/openwashdata/ploswater/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of ploswater is to make available data on publication trends from the [PLOS Water Journal](https://journals.plos.org/water/).

## Installation

You can install the development version of ploswater from
[GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("openwashdata/ploswater")
```

```{r}
## Run the following code in console if you don't have the packages
## install.packages(c("dplyr", "knitr", "readr", "stringr", "gt", "kableExtra"))
library(dplyr)
library(knitr)
library(readr)
library(stringr)
library(gt)
library(kableExtra)
```

Alternatively, you can download the individual datasets as a CSV or XLSX
file from the table below.

1.  Click Download CSV. A window opens that displays the CSV in
    your browser.
2.  Right-click anywhere inside the window and select "Save Page As...".
3.  Save the file in a folder of your choice.

```{r, echo=FALSE, message=FALSE, warning=FALSE}

extdata_path <- "https://github.com/openwashdata/ploswater/raw/main/inst/extdata/"

read_csv("data-raw/dictionary.csv") |> 
  distinct(file_name) |> 
  dplyr::mutate(file_name = str_remove(file_name, ".rda")) |> 
  dplyr::rename(dataset = file_name) |> 
  mutate(
    CSV = paste0("[Download CSV](", extdata_path, dataset, ".csv)"),
    XLSX = paste0("[Download XLSX](", extdata_path, dataset, ".xlsx)")
  ) |> 
  knitr::kable()

```

## Data

The package provides access to data from publications in the [PLOS Water Journal](https://journals.plos.org/water/). This data includes information on publication types, SDGs targeted, authors, institutions, and more.

```{r}
library(ploswater)
```

### ploswater

The dataset `ploswater` contains data about ... It has
`r nrow(ploswater)` observations and `r ncol(ploswater)`
variables

```{r}
ploswater |> 
  select(1:10) |>  # Select first 10 columns
  head(3) |> 
  gt::gt() |>
  gt::as_raw_html()
```

For an overview of the variable names, see the following table.

```{r echo=FALSE, message=FALSE, warning=FALSE}
readr::read_csv("data-raw/dictionary.csv") |>
  dplyr::filter(file_name == "ploswater.rda") |>
  dplyr::select(variable_name:description) |> 
  knitr::kable() |> 
  kableExtra::kable_styling("striped") |> 
  kableExtra::scroll_box(height = "200px")
```

## Example

```{r}
library(ploswater)
library(ggplot2)
library(lubridate)

# Convert publication_date to Date type
ploswater$publication_date <- as.Date(ploswater$publication_date)

# Create year-month column and count publications
monthly_counts <- ploswater %>%
  mutate(publication_month = floor_date(publication_date, "month")) %>%
  count(publication_month) %>%
  arrange(publication_month)

# Create bar chart
ggplot(monthly_counts, aes(x = publication_month, y = n)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(x = "",
       y = "Number of Publications",
       title = "Number of Publications per Month") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_x_date(date_breaks = "3 month", 
               date_labels = "%b %Y")
```

```{r}
# Create citation impact visualization
ggplot(ploswater, aes(x = publication_date, y = cited_by_count)) +
  geom_point(aes(size = fwci), alpha = 0.6) +
  scale_size_continuous(name = "Field-Weighted\nCitation Impact") +
  labs(x = "Publication Date",
       y = "Number of Citations",
       title = "Citation Impact Over Time") +
  theme_minimal()
```


```{r}
library(gt)

# Compute summary statistics
summary_stats <- data.frame(
  Statistic = c(
    "Mean Citations",
    "Median Citations",
    "Mean FWCI",
    "Total Publications",
    "Correlation: Citations & FWCI"
  ),
  Value = c(
    mean(ploswater$cited_by_count, na.rm = TRUE),
    median(ploswater$cited_by_count, na.rm = TRUE),
    mean(ploswater$fwci, na.rm = TRUE),
    round(nrow(ploswater), 0),
    cor(ploswater$cited_by_count, ploswater$fwci, use = "complete.obs")
  )
)

# Create and print the table using gt
summary_stats %>%
  gt() %>%
  tab_header(
    title = "Summary Statistics"
  )
```



## License

Data are available as
[CC-BY](https://github.com/openwashdata/%7B%7B%7Bpackagename%7D%7D%7D/blob/main/LICENSE.md).

## Citation

Please cite this package using:

```{r}
citation("ploswater")
```
